---
title: "BIOL343 Coding Challenge 2"
author: 'Group Number #'
date: "2026-09-07"
output: html_document
---

# BIOL343 Coding Challenge 2


## 1. Add your group members name with student id and the github repo
```{r group-members_with_the_github_repo}
# Name - Queen's Student Number (github repo)

# 1.

```

## 2. Run an setup chuck and when it is knitted it should not appear in the document.

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```
---

## 3. Now, import the dataset used in the coding challenge 1

```{r import-the-data}
cc1_data <-
```
---

## 4. Use functions you learnt previously from coding challenge 1 to inspect the data and retain the appropriate rows and columns. For this chunk, show only the code but not the printed output in the final knitted document.

```{r data-inspect}

```
---
## 5. Calculate summary statistics of average august temperature using `summary()` and construct an table. For this chunk, hide both code and results.

```{r summary-statistics}

```
##Summary statistics of average temperature in August


---
## 6. As a final step, insert an image of study species (i.e. *Solanum lycopersicum*) with the source of the image.

**INSERT HERE**

---


