---
title: "BIOL343 Coding Challenge 1"
author: "Group Members"
date: "2026-09-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BIOL343 Coding Challenge 1


## Group Members

```{r group-members}
# Name - Queen's Student Number

# 1.

```

# Part 1 — Import and Inspect the Dataset

## 1. Import the Data

The dataset is located in the `data` folder and is named:

`A1_Data_Excel.csv`

Use `read.csv()` to import the file and save it as an object called `assn1_data`.

```{r import-data}

assn1_data <- read.csv("")

```

---

## 2. Inspect the Data

Use the following functions to examine the dataset:

- `class()`
- `dim()`
- `str()`
- `head()`
- `tail()`

```{r inspect-data}

# class()


# dim()


# str()


# head()


# tail()

```

---

## 3. Clean the Dataset

Based on your inspection, retain the appropriate rows and columns.

Save the cleaned dataset as `assn1_data_mod`.

```{r clean-data}

assn1_data_mod <- 

```

---

## 4. Calculate Ovary Area

Create a new variable called `ovary_area` using:

\[
\text{ovary area} = \text{ovary width} \times \text{ovary height}
\]

```{r ovary-area}

assn1_data_mod$ovary_area <- 

```

Use `head()` to confirm that the new column was created.

```{r check-ovary-area}

# Check the first few rows.

```

# Part 2 — Simulate and Combine Data

You will now create a small simulated dataset with the same variables as the original dataset.

For continuous variables such as ovary width and ovary height, generate values from a **normal distribution** using `rnorm()`. Use a mean and standard deviation similar to those in the previously loaded `assn1_data_mod` dataset. You can calculate these values from the original data using `mean()` and `sd()`.

Before generating any random values, use `set.seed()` so that your simulated results can be reproduced.

Use the seed value `343`.

```{r set-seed}

set.seed(343)

```

> Run `set.seed(343)` before generating your simulated values. Using the same seed will produce the same random values each time the code is run.

## 5. Create Vectors

Create vectors for **10 new observations**.

### Population Type

Create a vector called `sim_population_type` containing `"A"` and `"S"` values.

```{r simulate-population}

sim_population_type <- 

```

### Ovary Width

Create a numeric vector called `sim_ovary_width` containing 10 simulated ovary-width values.

Use `rnorm()` and choose the mean and standard deviation based on the ovary-width values in `assn1_data_mod`.

```{r simulate-width}

#mean of ovary_width from assn1_data_mod


#sd of ovary_width from assn1_data_mod


sim_ovary_width <- 

```

### Ovary Height

Create a numeric vector called `sim_ovary_height` containing 10 simulated ovary-height values.

Use `rnorm()` and choose the mean and standard deviation based on the ovary-height values in `assn1_data_mod`.

```{r simulate-height}

#mean of ovary_height from assn1_data_mod


#sd of ovary_height from assn1_data_mod


sim_ovary_height <- 

```

### Flower Number

Create a numeric vector called `sim_flower_number` containing 10 simulated flower-count values.

```{r simulate-flowers}

#mean of flower_number from assn1_data_mod


#sd of flower_number from assn1_data_mod



sim_flower_number <- 

```

---

## 6. Create a Simulated Data Frame

Combine your four vectors into a data frame called `sim_data`.

Make sure the column names match the corresponding columns in `assn1_data_mod`.

```{r create-sim-data}

sim_data <- 

```

Calculate `ovary_area` for the simulated observations.

```{r simulated-ovary-area}

sim_data$ovary_area <- 

```

Use `str()` and `head()` to inspect your simulated dataset.

```{r inspect-sim-data}

# str()


# head()

```

---

## 7. Combine the Datasets

Combine the original cleaned dataset and the simulated dataset using `rbind()`.

Save the combined dataset as `combined_data`.

```{r combine-data}

combined_data <- 

```

Check the dimensions of the combined dataset.

```{r combined-dimensions}

# Check the dimensions.

```

# Part 3 — Basic Summary Statistics

## 8. Flower Number Summary Statistics

Calculate the following for `flower_number` in `combined_data`:

- Minimum
- Maximum
- Mean
- Median
- Standard deviation

```{r flower-summary}

# Minimum


# Maximum


# Mean


# Median


# Standard deviation

```

---

## 9. Use `summary()`

Use `summary()` on both `ovary_area` and `flower_number`.

```{r summary-function}

# Ovary area


# Flower number

```

**Question 1:** Which statistic that you calculated manually is not included in the output from `summary()`?

**Answer:**  

---

## 10. Compare Population Types

Create two subsets from `combined_data`:

- `asexual_data` for population type `"A"`
- `sexual_data` for population type `"S"`

```{r population-subsets}

asexual_data <- 


sexual_data <- 

```

Calculate the mean ovary area for each group.

```{r compare-populations}

# Mean ovary area for A


# Mean ovary area for S

```

**Question 2:** Which population type has the larger mean ovary area?

**Answer:**  

---

## 11. Visualization of ovary area

Use `ggplot2` to create a figure of your choice comparing ovary area between population types 

Your figure should include:

- appropriate axis labels
- a descriptive title
- legends if necessary

```{r load the library}

```

```{r visualization of ovary_width}

```
