---
title: "BIOL343 Coding Challenge 1"
author: "Group Members"
date: "2026-09-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BIOL343 Coding Challenge 1

## Group Members

```{r group-members}
# Name - Queen's Student Number

# 1. Sree (20376328)

```

# Part 1 — Import and Inspect the Dataset

## 1. Import the Data

The dataset is located in the `data` folder and is named:

`A1_Data_Excel.csv`

Use `read.csv()` to import the file and save it as an object called `assn1_data`.

```{r import-data}

assn1_data <- read.csv("data/A1_Data_Excel.csv")

```

---

## 2. Inspect the Data

Use the following functions to examine the dataset:

- `class()`
- `dim()`
- `str()`
- `head()`
- `tail()`

```{r inspect-data}

# class()
class(assn1_data)

# dim()
dim(assn1_data)

# str()

str(assn1_data)
# head()
head(assn1_data)

# tail()
tail(assn1_data)
```

---

## 3. Clean the Dataset

Based on your inspection, retain the appropriate rows and columns.

Save the cleaned dataset as `assn1_data_mod`.

```{r clean-data}

assn1_data_mod <- assn1_data[1:46, 1:8]

```

---

## 4. Calculate Ovary Area

Create a new variable called `ovary_area` using:

\[
\text{ovary area} = \text{ovary width} \times \text{ovary height}
\]

```{r ovary-area}

assn1_data_mod$ovary_area <- assn1_data_mod$ovary_width * assn1_data_mod$ovary_height

```

Use `head()` to confirm that the new column was created.

```{r check-ovary-area}

# Check the first few rows.
head(assn1_data_mod)
```

# Part 2 — Simulate and Combine Data

You will now create a small simulated dataset with the same variables as the original dataset.

For continuous variables such as ovary width and ovary height, generate values from a **normal distribution** using `rnorm()`. Use a mean and standard deviation similar to those in the previously loaded `assn1_data_mod` dataset. You can calculate these values from the original data using `mean()` and `sd()`.

Before generating any random values, use `set.seed()` so that your simulated results can be reproduced.

Use the seed value `343`.

```{r set-seed}

set.seed(343)

```

> Run `set.seed(343)` before generating your simulated values. Using the same seed will produce the same random values each time the code is run.

## 5. Create Vectors

Create vectors for **10 new observations**.

### Population Type

Create a vector called `sim_population_type` containing `"A"` and `"S"` values using `sample()` with size as `10` and replacement as`TRUE`

```{r simulate-population}

sim_population_type <- sample(
  assn1_data_mod$population_type,
  size = 10,
  replace = TRUE
)
```

### Ovary Width

Create a numeric vector called `sim_ovary_width` containing 10 simulated ovary-width values.

Use `rnorm()` and choose the mean and standard deviation based on the ovary-width values in `assn1_data_mod`.

```{r simulating ovary-width}

#calculate mean and sd of ovary-width

# mean
mean(assn1_data_mod$ovary_width, na.rm = TRUE)
# sd
sd(assn1_data_mod$ovary_width, na.rm = TRUE)


sim_ovary_width <- sim_ovary_width <- rnorm(
  n = 10,
  mean = 3.42,
  sd = 0.29)

```

### Ovary Height

Create a numeric vector called `sim_ovary_height` containing 10 simulated ovary-height values.

Use `rnorm()` and choose the mean and standard deviation based on the ovary-height values in `assn1_data_mod`.

```{r simulate-height}

#calculate mean and sd of ovary-height

# mean
mean(assn1_data_mod$ovary_height, na.rm = TRUE)
# sd
sd(assn1_data_mod$ovary_height, na.rm = TRUE)

sim_ovary_height <- rnorm(
  n = 10,
  mean = 3.025,
  sd = 0.3584395)

```

### Flower Number

Create a numeric vector called `sim_flower_number` containing 10 simulated flower-count values.

```{r simulate-flowers}
#calculate mean and sd of flower_number

# mean
mean(assn1_data_mod$flower_number, na.rm = TRUE)
# sd
sd(assn1_data_mod$flower_number, na.rm = TRUE)


sim_flower_number <- sim_ovary_height <- rnorm(
  n = 10,
  mean = 58.0156,
  sd = 44.13149)

```

---

## 6. Create a Simulated Data Frame

Combine your four vectors into a data frame called `sim_data`.

Make sure the column names match the corresponding columns in `assn1_data_mod`.

```{r create-sim-data}

sim_data <- data.frame(
  population_type = sim_population_type,
  ovary_width = sim_ovary_width,
  ovary_height = sim_ovary_height,
  flower_number = sim_flower_number
)

```

Calculate `ovary_area` for the simulated observations.

```{r simulated-ovary-area}

sim_data$ovary_area <- sim_data$ovary_width * sim_data$ovary_height


```

Use `str()` and `head()` to inspect your simulated dataset.

```{r inspect-sim-data}

# str()
str(sim_data)

# head()
head(sim_data)
```

---

## 7. Combine the Datasets

Combine the original cleaned dataset and the simulated dataset using `rbind()`.

Save the combined dataset as `combined_data`.

```{r combine-data}

combined_data <- rbind(
  assn1_data_mod[, names(sim_data)],
  sim_data
)

```

Check the dimensions of the combined dataset.

```{r combined-dimensions}

# Check the dimensions.
dim(combined_data)
```

# Part 3 — Basic Summary Statistics and Visualization


## 8. Flower Number Summary Statistics

Calculate the following for `flower_number` in `combined_data`:

- Minimum
- Maximum
- Mean
- Median
- Standard deviation

```{r flower-summary}

# Minimum
min(combined_data$flower_number, na.rm = TRUE)

# Maximum

max(combined_data$flower_number, na.rm = TRUE)
# Mean
mean(combined_data$flower_number, na.rm = TRUE)

# Median
median(combined_data$flower_number, na.rm = TRUE)

# Standard deviation
sd(combined_data$flower_number, na.rm = TRUE)
```

---

## 9. Use `summary()`

Use `summary()` on both `ovary_area` and `flower_number`.

```{r summary-function}

# Ovary area

summary(combined_data$ovary_area)
# Flower number
summary(combined_data$flower_number)
```

---

## 11. Compare Population Types

Create two subsets from `combined_data`:

- `asexual_data` for population type `"A"`
- `sexual_data` for population type `"S"`

```{r population-subsets}

asexual_data <- subset(
  combined_data,
  population_type == "A"
)


sexual_data <- subset(
  combined_data,
  population_type == "S"
)

```

Calculate the mean ovary area for each group.

```{r compare-populations}

# Mean ovary area for A
mean(asexual_data$ovary_area, na.rm = TRUE)

# Mean ovary area for S

mean(sexual_data$ovary_area, na.rm = TRUE)

```

**Question 1:** Which population type has the larger mean ovary area?

**Answer:**  

---

## 11. Visualization of ovary area

Use `ggplot2` to create a figure of your choice comparing ovary area between population types using `theme_classic`

Your figure should include:

- use `facet_wrap` or `facet_grid` if needed
- appropriate axis labels
- a descriptive title
- legends if necessary

```{r load the library}
library(ggplot2)
```

```{r visualization of ovary_width}

##boxplot is usually one for visualizing but they could also much scatter plot with facets or geom_histogram. Just make sure theme_classic() with appropriate axis, legends and a title is present.  
```
