---
title: "BIOL343 Coding Challenge 3"
author: "Group Members"
date: "2026"
output: html_document
---

# BIOL343 Coding Challenge 3

## Group Members

```{r group-members}
# Name - Queen's Student Number - Github repo
# 1.

```

# Part 1 — Import, Inspect, and Summarize

## 1. Import the Dataset

Import the provided dataset using `read.csv()` and save it as `floral_data`.

```{r import-data}

floral_data <- 

```

---

## 2. Inspect the Dataset

Use the following functions:

- `dim()`
- `head()`
- `tail()`
- `str()`
- `summary()`

```{r inspect-data}

# dim()


# head()


# tail()


# str()


# summary()

```

---

## 3. Calculate Basic Summary Statistics

For `face_diam_mm`, calculate the following separately for SI and SC plants:

- sample size
- mean
- median
- standard deviation

Use `na.rm = TRUE` where appropriate.

```{r summary-si-sc}

# SI plants


# SC plants

```

# Part 2 — Visualize SI and SC Populations

Load `ggplot2`.

```{r load-ggplot2}

```

## 4. Faceted Histogram

Create a histogram of `face_diam_mm`.

Your plot should:

- show separate panels for SI and SC plants using `facet_wrap()`
- use an appropriate `binwidth`
- use `theme_classic()`
- include appropriate axis labels with units
- include a descriptive title
- any warnings or messages generated by code chunks should be hidden in the final knitted document using appropriate R Markdown chunk options.


```{r face-diameter-histogram}

# Create your histogram here.

```

---

## 5. Figure for Umbel Angle by Mating System

Create a appropriate figure comparing `umbel_angle_deg` between SI and SC plants.

Your plot should:

- use `facet_wrap()` if needed
- use `theme_classic()` for theme
- include axis labels, appropriate title, and captions
- the code chunk should be hidden and only figure and captions should be present in the final knitted document.
- any warnings or messages generated by code chunks should be hidden in the final knitted document

```{r umbel-angle-figure}

# Create your figure here.

```

# Part 3 — Relationship Between Floral Traits

## 6. Floral Face Diameter vs. Tube Length

Create a figure showing the relationship between:

- `face_diam_mm`
- `tube_length_mm`

Your plot should:

- colour points by `self_inc`
- add a separate linear regression line for each mating system with standard error
- use `theme_classic()`
- include axis labels with units, appropriate title, legends and captions
- the code chunk should be hidden and only figure and captions should be present in the final knitted document.
- any warnings or messages generated by code chunks should be hidden in the final knitted document

```{r floral-trait-figure}

# Create your figure here.

```

---

## 7. Interpretation

Explain your interpretation based on the figures generated above.

