---
title: "BIOL343 Coding Challenge 3"
author: "Group Members"
date: '2026'
output:
  pdf_document: default
  html_document: default
---

# BIOL343 Coding Challenge 3

## Group Members

```{r group-members}
# Name - Queen's Student Number - Github repo
# 1. Sree - 20376328 (https://github.com/sreevatshan/BIOL343_F26)

```

# Part 1 — Import, Inspect, and Summarize

## 1. Import the Dataset

Import the provided dataset using `read.csv()` and save it as `floral_data`.

```{r import-data}

floral_data <- read.csv("data/AbroniaFloralData.csv")

```

---

## 2. Inspect the Dataset

Use the following functions:

- `dim()`
- `head()`
- `tail()`
- `str()`

```{r inspect-data}

# dim()

dim(floral_data)

# head()
 head(floral_data)

# tail()

tail(floral_data)
# str()

str(floral_data)

```

---

## 3. Calculate Basic Summary Statistics

For `face_diam_mm`, calculate the following separately for SI and SC plants:

- sample size
- mean
- median
- standard deviation

Use `na.rm = TRUE` where appropriate.

```{r summary-si-sc}

SI <- subset(floral_data, self_inc == "SI")
SC <- subset(floral_data, self_inc == "SC")

# SI plants
summary(SI)
sd(SI$face_diam_mm)


# SC plants
summary(SC)
sd(SC$face_diam_mm)

```

# Part 2 — Visualize SI and SC Populations

Load `ggplot2`.

```{r load-ggplot2}
library(ggplot2)
```

## 4. Faceted Histogram

Create a histogram of `face_diam_mm`.

Your plot should:

- show separate panels for SI and SC plants using `facet_wrap()`
- use an appropriate `binwidth`
- use `theme_classic()`
- include appropriate axis labels with units
- include a descriptive title
- Any warnings or messages generated by code chunks should be hidden in the final knitted document using appropriate R Markdown chunk options.

```{r face-diameter-histogram}

# Create your histogram here.

ggplot(floral_data, aes(x= face_diam_mm)) + geom_histogram() + facet_wrap(~ self_inc) + theme_classic() + labs(
    x = "Floral face diameter (mm)",
    y = "Frequency",
    title = "Distribution of floral face diameter by mating system"
  )

```

---

## 5. Figure for Umbel Angle by Mating System

Create a appropriate figure comparing `umbel_angle_deg` between SI and SC plants.

Your plot should:

- use `facet_wrap()` if needed
- use `theme_classic()` for theme
- include axis labels, appropriate title, and captions
- the code chunk should be hidden and only figure and captions should be present in the final knitted document.
- any warnings or messages generated by code chunks should be hidden in the final knitted document
```{r umbel-angle-figure}

ggplot(floral_data,aes(x = self_inc,y = umbel_angle_deg)) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Mating system",y = "Umbel angle (degrees)",title = "Umbel angle by mating system")

```

# Part 3 — Relationship Between Floral Traits

## 6. Floral Face Diameter vs. Tube Length

Create a figure showing the relationship between:

- `face_diam_mm`
- `tube_length_mm`

Your plot should:

- colour points by `self_inc`
- add a separate linear regression line for each mating system with standard error
- use `theme_classic()`
- include axis labels with units, appropriate title, legends and captions
- the code chunk should be hidden and only figure and captions should be present in the final knitted document.
- any warnings or messages generated by code chunks should be hidden in the final knitted document

```{r floral-trait-figure}

# Create your figure here.

ggplot(floral_data,
  aes(x = face_diam_mm,y = tube_length_mm,colour = self_inc)) +
  geom_point() +
  geom_smooth(method = "lm",se = TRUE) +
  theme_classic() +
  labs(x = "Floral face diameter (mm)",y = "Floral tube length (mm)",colour = "Mating system",
    title = "Relationship between floral face diameter and tube length")

```

---

## 7. Interpretation

Explain your interpretation based on the figures generated above.

Student should say SC plants have less investment on floral traits which has association to pollination attraction. 

